% =========================================================
% Design Specifications and Material Selection
% =========================================================

\section{Design Specifications and Material Selection}
\label{sec:design-specifications}

The public portfolio edition treats the operating frequency, substrate, and laminate thickness as
fixed engineering design inputs. These parameters define the analytical patch synthesis, CST
material model, frequency-domain simulation range, farfield-monitor frequency, and array-spacing
calculations used throughout the project.

\subsection{Operating Frequency}

The selected design frequency is
\[
f_0 = 2.501~\text{GHz}.
\]
This value is used consistently as the reference frequency for the analytical calculations, CST
tuning process, farfield evaluation, and beam-steering analysis.

The corresponding free-space wavelength is
\[
\lambda_0
=
\frac{c}{f_0}
=
119.869~\text{mm},
\]
where \(c\) is the speed of light in free space.

\subsection{Substrate and Laminate Thickness}

The antenna is designed using DiClad~880-IM, a low-loss microwave laminate suitable for printed
RF and antenna structures. The nominal dielectric properties adopted for the analytical calculations
and CST material model are based on the manufacturer's published laminate data
\cite{rogers_diclad870_880}:
\[
\varepsilon_r = 2.20,
\qquad
\tan\delta = 0.0009.
\]
The selected laminate thickness is
\[
h = 1.524~\text{mm} = 0.060~\text{in}.
\]

The relative permittivity and loss tangent are treated as nominal design values. In a fabricated
implementation, manufacturing tolerance, material-lot variation, copper roughness, and frequency
dependence of the laminate properties could introduce additional resonance and efficiency variation.

Table~\ref{tab:design-specifications} summarizes the principal design inputs used throughout the
report.

\begin{table}[htbp]
    \centering
    \caption{Principal antenna design specifications and material parameters.}
    \label{tab:design-specifications}
    \small
    \renewcommand{\arraystretch}{1.15}
    \begin{tabularx}{0.90\textwidth}{@{}>{\raggedright\arraybackslash}X
        >{\raggedright\arraybackslash}p{0.40\textwidth}@{}}
        \toprule
        \textbf{Design Parameter} & \textbf{Selected Value} \\
        \midrule
        Design frequency, \(f_0\) & \(2.501~\text{GHz}\) \\
        Free-space wavelength, \(\lambda_0\) & \(119.869~\text{mm}\) \\
        Substrate & DiClad~880-IM \\
        Substrate thickness, \(h\) & \(1.524~\text{mm}\) \((0.060~\text{in})\) \\
        Relative permittivity, \(\varepsilon_r\) & \(2.20\) \\
        Loss tangent, \(\tan\delta\) & \(0.0009\) \\
        Single-element feed topology & Inset-fed microstrip line \\
        Nominal feed impedance & \(50~\Omega\) \\
        Array topology & Two-port \(2\times1\) linear patch array \\
        Initial array spacing & \(d \approx \lambda_0/2 = 59.935~\text{mm}\) \\
        \bottomrule
    \end{tabularx}
\end{table}
